A sample computation

### 2024-10-02 draft_systoles.tex
added the stiefel_whitney class

\section{Intro}
In this note we establish the following results:
\begin{enumerate}
\item
The regularor of $\tilde A$
grows as $\approx 3 \ln^2(p)$.

@@

\section{From $\SL$ to $\SLn$}

>>>>

\section{Intro}
In this note we establish the following results:
\begin{enumerate}
\item
The regulator of $\tilde A$
grows as $\approx 3 \ln^2(p)$.

@@

\newpage
\section{Computing (co)homology}
Christophe Soulé \cite{soule1978}
computed the cohomology of $\SL$
to be
\begin{table}[h!]
	\centering
\begin{tabular}{|c|c|c|c|c|c|c}
	\hline
	$n$        & $0$                               & $1$                               & $2$                               & $3$                               & $4$  & \ldots\\
\hline 
	$H^n(\SL, \mathbb{Z})$ & $\mathbb{Z}$ & $0$ & $0$ & $(\mathbb{Z}/2\mathbb{Z})^{2} $ & $(\mathbb{Z}/3\mathbb{Z})^{2} \oplus (\mathbb{Z}/4\mathbb{Z})^{2} $ & \ldots \\
\hline
\end{tabular}.
\end{table}

The universal coefficient formula gives us the homology.
\begin{table}[h!]
	\centering
\begin{tabular}{|c|c|c|c|c|c}
	\hline
	$n$        & $0$                               & $1$                               & $2$                               & $3$                               & \ldots\\
\hline 
	$H_n(\SL, \mathbb{Z})$ & $\mathbb{Z}$ & $0$ &  $(\mathbb{Z}/2\mathbb{Z})^{2} $ & $(\mathbb{Z}/3\mathbb{Z})^{2} \oplus (\mathbb{Z}/4\mathbb{Z})^{2} $ & \ldots \\
\hline
\end{tabular}
\end{table}

Recall the matrices 
\[
A = 
\begin{pmatrix}
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & -p^2
\end{pmatrix}
,
B = \Id + p A =
\begin{pmatrix}
1 & 0 & p \\
p & 1 & 0 \\
0 & p & 1 - p^3
\end{pmatrix}
.\] 
We denote $\tilde A$ by $B$ for convenience.

As $A, B $ commute they define a map
\[
f: \mathbb{Z}^2 \to \SL, 
\] 
\[
(k, l) \mapsto A^k B^l
.\] 
\fbox{
\parbox{\textwidth}{
The goal for us it to understand the map 
\[
H_2(\mathbb{Z}^2, \mathbb{Z}) \cong \mathbb{Z} \to 
H_2(\SL, \mathbb{Z}) \cong (\mathbb{Z}/2\mathbb{Z})^{2} 
.\] 
}}

Consider the natural embedding 
\[
i: \SL \embeds {\mathbf{SL}_3(\mathbb{R})}
.\] 
\begin{Prop}
The pullback 
\[
i^*: H^2(B\mathbf{SL}_3(\mathbb{R}), \mathbb{Z}/ 2 \mathbb{Z}) \cong \mathbb{Z} / 2 \mathbb{Z} \to 
H^2(\SL, \mathbb{Z}/ 2 \mathbb{Z}) \cong (\mathbb{Z}/2\mathbb{Z})^{2}
\] 
is an embedding.
\end{Prop}
\begin{proof}
Consider a pair of matrices
\[
C_1 = 
\begin{pmatrix}
-1 & 0 & 0 \\
0 & -1 & 0 \\
0 & 0 & 1
\end{pmatrix}
,
C_2 =
\begin{pmatrix}
1 & 0 & 0 \\
0 & -1 & 0 \\
0 & 0 & -1
\end{pmatrix} \in \SL
.\] 
They give rise to a principle $\SL$ bundle $E$ on $\mathbb{T}^2$ with non-zero $w_2$, the second Stiefel-Whitney class.

Indeed, by the splitting principle we have
\[
w(E) = w(E_1) w(E_2) w (E_3)
= (1 + x) (1 + x + y) (1 + y) 
= 1 
+ 2 (x + y) 
+ (x^2 + xy + xy + xy + y^2) 
+ x ( x + y) y 
= 1 + xy
,\] 
where $x, y$ denote the generators of $H^1(\mathbb{T}^2, \mathbb{Z}/ 2 \mathbb{Z})$.
Therefore the composition 
\[
H^2(B\mathbf{SL}_3(\mathbb{R}), \mathbb{Z}/ 2 \mathbb{Z}) \cong \mathbb{Z} / 2 \mathbb{Z} \to 
H^2(\SL, \mathbb{Z}/ 2 \mathbb{Z}) \cong (\mathbb{Z}/2\mathbb{Z})^{2}
\to
H^2(\mathbb{T}^2, \mathbb{Z}/ 2 \mathbb{Z}) \cong (\mathbb{Z}/2\mathbb{Z})
\] 
is nonzero and thus the first map is an embedding.
\end{proof}
Thus $w_2$ is one of the generators of the second cohomology of $\SL$.
Consider the principal $\SL$-bundle $E$ on $\mathbb{T}^2$ associated with $A, B$.
\begin{Prop}
	$w_2(E) = 0$.
\end{Prop}
\begin{proof}
By the splitting principle
\[
w(E) = w(E_1) w(E_2) w (E_3)
= (1 + x + y) (1 + x + y) (1) 
= 1
.\]
\end{proof}
There are perfect pairings
\[
H_2(\SL, \mathbb{Z}) \times 
H^2(\SL, \mathbb{Z}/2 \mathbb{Z}) \to \mathbb{Z}/ 2 \mathbb{Z}
,\] 
\[
H_2(\mathbb{T}^2, \mathbb{Z}) \times 
H^2(\mathbb{T}^2, \mathbb{Z}/2 \mathbb{Z}) \to \mathbb{Z}/ 2 \mathbb{Z}
.\] 
Combining them, we get that the map
\[
H_2(\mathbb{Z}^2, \mathbb{Z}) \cong \mathbb{Z} \to 
H_2(\SL, \mathbb{Z}) \cong (\mathbb{Z}/2\mathbb{Z})^{2} 
\] 
has one coordinate equal to zero.

\newpage
\section{From $\SL$ to $\SLn$}
